---
title: "Smoking mother and their baby’s birth weight "
Author: Zehui Zhang, Lu Xu
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

# 2. Basic Analysis
### 2.1 Data Processing & Numerically Analysis
Break the original data set into two subsets, smoked one and non-smoked one.
Remove the invalid data, where shows 9 in smoke column.
It seems pretty direct that, here, 0 means nonsmoke, and 1 means smoke. So 9 appared here is invalid data.
```{r}
# Clean the data, and make it into 2 sub-data sets(smoked, nonsmoked)
setwd("~/Downloads")
Ori_data = read.table('babies.txt', header=T)
```
Smoke_data
```{r}
smoke_ind = which(Ori_data$smoke > 0 & Ori_data$smoke < 9)
smoke_data = Ori_data[smoke_ind,]
```
Nonsmoke_data
```{r}
nonsmoke_ind = which(Ori_data$smoke < 1)
nonsmoke_data = Ori_data[nonsmoke_ind,]
```
Find some special values, like mean, mode, and standard deviation, etc.
```{r}
summary(smoke_data)
summary(nonsmoke_data)

smk = smoke_data$bwt
nsmk = nonsmoke_data$bwt

smk_sd = sd(smk)
nsmk_sd = sd(nsmk)
smk_sd
nsmk_sd

smk_bar = mean(smk)
nsmk_bar = mean(smk)
smk_bar
nsmk_bar

#helper function that help to find the mode from a data set
getmode <- function(v) {
  uniqv <- unique(v)
  uniqv[which.max(tabulate(match(v, uniqv)))]
  }
smk_mode = getmode(smk)
nsmk_mode = getmode(nsmk)
smk_mode
nsmk_mode
```
Standardize the data
```{r}
std_smk = (smk - smk_bar)/smk_sd
std_nsmk = (nsmk - nsmk_bar)/nsmk_sd
```
To check the normality by 68-95-99.7 rule
```{r}
# Check the normality of smoked one
Fs = ecdf(std_smk)
Fs(3) - Fs(-3)
Fs(2) - Fs(-2)
Fs(1) - Fs(-1)

# Check the normality of nonsmoked one
Fn = ecdf(std_nsmk)
Fn(3) - Fn(-3)
Fn(2) - Fn(-2)
Fn(1) - Fn(-1)
```
To check the normality by skewness and kurtosis
```{r}
# Function to find the Skewness Coefficient
skew = function(X) {
  Xbar = mean(X); S = sd(X)
  Z = (X - Xbar)/S
  mean(Z^3)
}

skew(std_smk)
skew(std_nsmk)

#Function to find the Kurtosis Coefficient
kurt = function(X) {
  Xbar = mean(X); S = sd(X)
  Z = (X - Xbar)/S
  mean(Z^4)
}

kurt(std_smk)
kurt(std_nsmk)
```
By using the foor loop to find the skewness and kurtosis coefficient of a normal distribution with 742 observations by 10000 times.
```{r}
S1 = NULL
for(i in 1:10000){
  S1[i] = skew(rnorm(484))
}
summary(S1)

K1 = NULL
for(i in 1:10000){
  K1[i] = kurt(rnorm(484))
}
summary(K1)

S2 = NULL
for(i in 1:10000){
  S2[i] = skew(rnorm(742))
}
summary(S2)

K2 = NULL
for(i in 1:10000){
  K2[i] = kurt(rnorm(742))
}
summary(K2)
```
### 2.2 Use graphical methods to compare the two distributions of birth weight.
Create the histograms
```{r}
# histograms for smoked group
hist(smk, breaks = 30, probability=T)
curve(dnorm(x, mean = 114.1, sd = 18.1), 60, 160, col='red', add=T)

# histograms for nonsmoked group
hist(nsmk, breaks = 30, probability = T)
curve(dnorm(x, mean = 123, sd = 17.4), 60, 180, col='red', add=T)
```
Check the normality of two sub data sets
```{r}
# QQ plot(Quantile-Quantile Plot) for smoked group
qqnorm(std_smk)
abline(0, 1, col='red', lwd=2)

# QQ plot(Quantile-Quantile Plot) for nonsmoked group
qqnorm(std_nsmk)
abline(0, 1, col='red', lwd=2)
```
### 2.3 Comparison of incidence
Find the proportion of the smoked mothers and non smoked.
```{r}
# % smoked
length(smk)/length(Ori_data$bwt) *100
# % nonsmoked
length(nsmk)/length(Ori_data$bwt) *100
```
Find the incidence of low-birth-weight when low-birth-weight is below 88
```{r}
LBW_smk1 = length(which(smk < 88))
LBW_nsmk1 = length(which(nsmk < 88))

# incidence for smoker
smk_i1 = LBW_smk1 / length(smk)
# incidence for nonsmoker 
nsmk_i1 = LBW_nsmk1 / length(nsmk)
smk_i1
nsmk_i1
LBW_smk1
LBW_nsmk1
```
Find the incidence of low-birth-weight when low-birth-weight is below 60
```{r}
LBW_smk2 = length(which(smk < 60))
LBW_nsmk2 = length(which(nsmk < 60))

# incidence for smoker
smk_i2 = LBW_smk2 / length(smk)
# incidence for nonsmoker 
nsmk_i2 = LBW_nsmk2 / length(nsmk)
smk_i2
nsmk_i2
LBW_smk2
LBW_nsmk2
```
Find the incidence of low-birth-weight when low-birth-weight is 100
```{r}
LBW_smk3 = length(which(smk < 100))
LBW_nsmk3 = length(which(nsmk < 100))

# incidence for smoker
smk_i3 = LBW_smk3 / length(smk)
# incidence for nonsmoker 
nsmk_i3 = LBW_nsmk3 / length(nsmk)
smk_i3
nsmk_i3
LBW_smk3
LBW_nsmk3
```
Showing these value in the histogram
```{r}
# for smoker
hist(smk)
abline(v = 88, col='green')
abline(v = 60, col='blue')
abline(v = 100, col='red')

# for nonsmoker
hist(nsmk)
abline(v = 88, col='green')
abline(v = 60, col='blue')
abline(v = 100, col='red')
```
Measure the difference:
```{r}
# difference when LBW is 88
smk_i1 - nsmk_i1
# difference when LBW is 60
smk_i2 - nsmk_i2
# difference when LBW is 100
smk_i3 - nsmk_i3
```
# Advanced Analysis
Seperate into 3 groups
```{r}
# Age between 18 - 30
young_ind = which(Ori_data$age < 30 & Ori_data$age >= 18)
young = Ori_data[young_ind,]

# Age between 31-40
middle_ind = which(Ori_data$age < 40 & Ori_data$age >= 31)
middle = Ori_data[middle_ind,]

# Age greater than 41
old_ind = which(Ori_data$age > 40)
old = Ori_data[old_ind,]
```
Find mean, median, mode, and incidence of low-birth-weight
```{r}
# find mean, median, and mode
summary(young)
summary(middle)
summary(old)
```
```{r}
y_mode = getmode(young$bwt)
m_mode = getmode(middle$bwt)
o_mode = getmode(old$bwt)
y_mode
m_mode
o_mode
```
```{r}
# find incidence
LBW_young = length(which(young$bwt < 88))
LBW_middle = length(which(middle$bwt < 88))
LBW_old = length(which(old$bwt < 88))

young_incidence = LBW_young / length(young$bwt)
middle_incidence = LBW_middle / length(middle$bwt)
old_incidence = LBW_old / length(old$bwt)

young_incidence
middle_incidence
old_incidence
LBW_young
LBW_middle
LBW_old
```
Find the skewness
```{r}
skew(young$bwt)
skew(middle$bwt)
skew(old$bwt)
```
