---
title: "Snow Gauge Calibration"
author: Lu Xu and Zehui Zhang
output: pdf_document

---

# 2. Basic Analysis
So as to determine a deoth profile of snow density, in this case, we will develop a procedure to calibrate the snow gauge from data

Open the data set
```{r}
setwd("~/Downloads")
dat <- read.table("gauge.txt", header=TRUE)
```

### 2.1 Raw Data
By using R, fit this data set to a liner regression model, and plot it out. Then check if the transformation needs. 

Make the sactterplot and the best fit linear regression line
```{r}
# sactter plot and the linear regression model
plot(dat$gain ~ dat$density, ylab = "Gain", xlab = "Density",pch = 19)
Lm = lm(dat$gain ~ dat$density, data = dat)
abline(Lm,col = 'red',lwd = 3)
summary(Lm)
```
Make the residual plot
```{r}
plot(residuals(Lm) ~ dat$gain, ylab = "Residual", xlab = 'Density',pch =19)
abline(h = 0,col = 'red',lwd = 3)
```
By making the histogram and the qq plot of the residual, it is efficient to check if the transformation needed. 
```{r}
hist(Lm$residuals)
qqnorm(Lm$residuals,pch = 19)
qqline(Lm$residuals,col = "red")
```

### 2.2 Transformation of Data
Based on the work in the 2.1 section, we found that the transformation is necessary. In this part, we will transform the data into an appropriate model and justify this model is appropriate one. 

Log Transformation
```{r}
gain = dat$gain
density = dat$density
Logm = lm(log(gain) ~ density, data = dat)
summary(Logm)
```
Make the scatter plot after transformation
```{r}
plot(log(gain) ~ density, data = dat, pch = 16) ; grid(lty='solid')
lines(predict(Logm) ~ density, data = dat, pch = 16, col = "red")
```
Plot the residual plot
```{r}
plot(residuals(Logm) ~ density, data = dat, pch = 16)
abline(h = 0, lwd = 2, col = "red")
```
Create the histogram and the pp plot of the residual, so as to check if this model appropriate
```{r}
hist(residuals(Logm))
qqnorm(Logm$residuals,pch = 19)
qqline(Logm$residuals,col = "red")
```

### 2.3 Robustness
Due to the densities of the polyethylene blocks are not reported exactly, we will do the simulation study by add a random error into the data set for both linear model and the transformed model.

Add the random error to the original density
Data set 1
```{r}
new_dat1 = jitter(density, factor = 1, amount = NULL)
dat$new_data1 = new_dat1
ndensity1 = dat$new_data1
```
Data set 2
```{r}
new_dat2 = jitter(density, factor = 1, amount = NULL)
dat$new_data2 = new_dat2
ndensity2 = dat$new_data2
```
Simulation study for the linear model
```{r}
# For data set 1
plot(dat$gain ~ ndensity1, ylab = "Gain", xlab = "New Density 1",pch = 19)
nLm1 = lm(dat$gain ~ ndensity1, data = dat)
abline(nLm1,col = 'red',lwd = 3)
summary(nLm1)

plot(residuals(nLm1) ~ dat$gain, ylab = "Residual", xlab = 'New Density 1',pch =19)
abline(h = 0,col = 'red',lwd = 3)

hist(nLm1$residuals)
qqnorm(nLm1$residuals,pch = 19)
qqline(nLm1$residuals,col = "red")
```
```{r}
# for data set 2
plot(dat$gain ~ ndensity2, ylab = "Gain", xlab = "New Density 2",pch = 19)
nLm2 = lm(dat$gain ~ ndensity2, data = dat)
abline(nLm2,col = 'red',lwd = 3)
summary(nLm2)

plot(residuals(nLm2) ~ dat$gain, ylab = "Residual", xlab = 'New Density 2',pch =19)
abline(h = 0,col = 'red',lwd = 3)

hist(nLm2$residuals)
qqnorm(nLm2$residuals,pch = 19)
qqline(nLm2$residuals,col = "red")
```
Simulation study for log model
```{r}
# For Data Set 1
nLogm1 = lm(log(gain) ~ ndensity1, data = dat)
summary(nLogm1)
plot(log(gain) ~ ndensity1, data = dat, pch = 16) ; grid(lty='solid')
lines(predict(nLogm1) ~ ndensity1, data = dat, pch = 16, col = "red")
plot(residuals(nLogm1) ~ ndensity1, data = dat, pch = 16)
abline(h = 0, lwd = 2, col = "red")
hist(residuals(nLogm1))
qqnorm(nLogm1$residuals,pch = 19)
qqline(nLogm1$residuals,col = "red")
```
```{r}
# For Data Set 2
nLogm2 = lm(log(gain) ~ ndensity2, data = dat)
summary(nLogm2)
plot(log(gain) ~ ndensity2, data = dat, pch = 16) ; grid(lty='solid')
lines(predict(nLogm2) ~ ndensity2, data = dat, pch = 16, col = "red")
plot(residuals(nLogm2) ~ ndensity2, data = dat, pch = 16)
abline(h = 0, lwd = 2, col = "red")
hist(residuals(nLogm2))
qqnorm(nLogm2$residuals,pch = 19)
qqline(nLogm2$residuals,col = "red")
```

### 2.4 Forward prediction
Produce point estimates and uncertainty bands for gains, specifically, 0.508, 0.001
```{r}
# Prediction bounds
p1<-predict(Logm,newdata = data.frame(density = unique(density)), interval ="prediction")
p1
dat2 <- data.frame(density = density,log = log(gain))
plot(dat2,pch = 19)
abline(Logm)
lines(p1[,2]~ unique(density),lty = 2,col = "blue")
lines(p1[,3]~ unique(density),lty = 2,col = "blue")

#0.508 
pred1 = predict(Logm,newdata = data.frame(density = 0.508),interval = "prediction")
exp(pred1)

#0.001
pred2 = predict(Logm,newdata = data.frame(density = 0.001),interval = "prediction")
exp(pred2)
```
### 2.5 Reverse prediction prediction
Invert the forward prediction made in previous part, and check the corresponding gain of the real density value. 
```{r}
# Invert model
iLogm = lm(log(density) ~ gain, data = dat)
summary(iLogm)
```
```{r}
newdensity = data.frame(gain = c(38.6,426.7))
p2 <- predict(iLogm,newdata = newdensity, interval ="prediction")
exp(p2)
dat3 <- data.frame(gain = gain, density = log(density))
plot(dat3,pch = 19)
abline(iLogm)
lines(p2[,2]~ newdensity$gain,lty = 2,col = "blue")
lines(p2[,3]~ newdensity$gain,lty = 2,col = "blue")
```

### 2.6 Corss-Validation 
SO as to avoid the condition that the measurement corresponding to the densities 0.508 and 0.001 were included in the fitting. Thus in this part, we will remove these data from the original data set then make a new predict. 

Remove the 0.508 and 0.001
```{r}
setwd("~/Downloads")
dat <- read.table("gauge.txt", header=TRUE)
sub1 = subset(dat, dat$density != 0.508)
sub2 = subset(dat, dat$density != 0.001)
```
prediction for the subset 1
```{r}
# create the model 
sub1density = sub1$density
sub1gain = sub1$gain
sub1_iLogm = lm(log(sub1density) ~ sub1gain, data = sub1)
summary(sub1_iLogm)

# make the prediction
newset = data.frame(sub1gain = c(38.6))
sub1p <- predict(sub1_iLogm, newdata = newset, interval ="prediction")
exp(sub1p)
```
prediction for subset2
```{r}
# create the model 
sub2density = sub2$density
sub2gain = sub2$gain
sub2_iLogm = lm(log(sub2density) ~ sub2gain, data = sub2)
summary(sub2_iLogm)

# make the prediction
newset = data.frame(sub2gain = c(38.6))
sub2p <- predict(sub2_iLogm, newdata = newset, interval ="prediction")
exp(sub2p)
```

# 3 Advanced Analysis
In the precious analysis, we only used the log transformation, but is there some better models? In this part, we will use cross-validation to find if polynomial regression works here. Specifically, what quadratic regression and cubic regression behave compared with log regression
```{r}
# Quadratic model
m2 <- lm(density ~ gain + I(gain^2), data=dat)
plot(density ~ gain, data=CO2, pch=16); grid(lty='solid')
lines(predict(m2) ~ gain, data=dat, col='red', lwd=2)
plot(residuals(m2) ~ gain, data=dat, pch=16)
abline(h = 0, lwd = 2)
summary(m2)

# make the prediction
newset_m2 = data.frame(gain = c(38.6))
predict(m2, newdata = newset_m2, interval ="prediction")

# residual between best expect and the real data
abs(0.53064383 - 0.508)
```
```{r}
# Cubic regression
m3 <- lm(density ~ gain + I(gain^2) + I(gain^3), data=dat)
plot(density ~ gain, data=dat, pch=16); grid(lty='solid')
lines(predict(m3) ~ gain, data=dat, col='red', lwd=2)
plot(residuals(m3) ~ gain, data=dat, pch=16)
abline(h = 0, lwd = 2)
summary(m3)

# make the prediction
newset_m3 = data.frame(gain = c(38.6))
predict(m3, newdata = newset_m3, interval ="prediction")

# residual between best expect and the real data
abs(0.532838748 - 0.508)
```
The residual between real value and the best fit for log model
```{r}
0.654081341 - 0.508
```

